
#Loading the datasets

```{r}
path = "~/Library/Mobile Documents/com~apple~CloudDocs/SMU MITB/ISSS616 - Applied Statistical Analysis with R/Group Project/"
happy = read.csv(paste0(path, "Chapter2OnlineData.csv"))
countries = read.csv(paste0(path, "countries of the world.csv"), dec=',')
world = read.csv(paste0(path, "world.csv"), na.string='..')
head(happy)
head(countries)
head(world)
```

#Checking the dimension of the data

```{r}
print(dim(happy))
print(dim(countries))
print(dim(world))
```

#Checking happiness dataset (overall)

```{r}
str(happy)
```

#Checking happiness dataset (NA value)
```{r}
par(mar=c(10,4,4,4))
plt <- barplot(sapply(happy, function(x) sum(is.na(x))), xlab="", col='steelblue', xaxt="n",
               ylab = "frequency")
text(plt, par("usr")[3], labels=colnames(happy), srt=60, xpd=T, cex=0.5, adj=1)
title("NA in each column")
mtext(side=1, text="columns", line=9)
```

# Remove columns with lots of missing data and useless columns [15:17], [19:]

```{r}
happy = happy[-c(15:17, 19:26)]
colnames(happy)
```


# Calculate the mean values for all attributes from year 2011 to 2018
# remove all columns except country, attribute and average
# further remove garbage rows at the end

```{r}
world$average = rowMeans(world[,5:12], na.rm=TRUE)
world = world[c(1, 3, 13)]
world = world[1:3689,]
```

# Convert the long format to short format
```{r}
library(tidyverse)
world = world %>%
  spread(Series.Name, average)
```

#Checking countries dataset for missing data
```{r}
sapply(world, function(x) sum(is.na(x)))
```

# What are the two countries with missing 'Land area (sq. km)'?
```{r}
world[which(is.na(world['Land area (sq. km)'])),]
```

# Remove columns with more than 25 missing countries
```{r}
world = world[colnames(world)[sapply(world, function(x) sum(is.na(x))) < 25]]
```

# Merging happiness and world dataset
```{r}
new_table = left_join(x=happy, y=world, by=c("Country.name" = "Country.Name") )
head(new_table)
```

# Examine the new table - countries that are not matched
```{r}
unique(new_table[is.na(new_table['Land area (sq. km)']),][['Country.name']])
```

# Need to change the names manually
Congo (Brazzaville) -> Congo, Repub. of the 
Congo (Kinshasa) -> Congo, Dem. Rep
Egypt -> Egypt, Arab Rep.
Gambia -> Gambia, The
Hong Kong S.A.R. of China -> Hong Kong SAR, China
Iran -> Iran, Islamic Rep.
Ivory Coast -> Cote d'Ivoire
Kyrgyzstan -> Kyrgyz Republic
Laos -> Lao PDR
Macedonia -> North Macedonia
North Cyprus -> Cyprus
Palestinian Territories -> West Bank and Gaza
Russia -> Russian Federation
Slovakia -> Slovak Republic
Somaliland region -> Somalia
South Korea -> Korea, Rep.
Swaziland -> Eswatini
Syria -> Syrian Arab Republic
Taiwan Province of China -> ***MISSING***
Venezuela -> Venezuela, RB
Yemen -> Yemen, Rep.


```{r}
## in happy dataset
from_c = c('Congo (Brazzaville)', 'Congo (Kinshasa)', 'Egypt', 'Gambia',
           'Hong Kong S.A.R. of China', 'Iran', 'Ivory Coast', 'Kyrgyzstan',
           'Laos', 'Macedonia', 'North Cyprus', 'Palestinian Territories',
           'Russia', 'Slovakia', 'South Korea', 'Swaziland', 'Syria',
           'Venezuela', 'Yemen', 'Somaliland region')
## in world dataset
to_c = c('Congo, Rep.', 'Congo, Dem. Rep.', 'Egypt, Arab Rep.', 'Gambia, The',
         'Hong Kong SAR, China', 'Iran, Islamic Rep.', 'Cote d\'Ivoire', 'Kyrgyz Republic',
         'Lao PDR', 'North Macedonia', 'Cyprus', 'West Bank and Gaza',
         'Russian Federation', 'Slovak Republic', 'Korea, Rep.',
         'Eswatini', 'Syrian Arab Republic', 'Venezuela, RB', 'Yemen, Rep.', 'Somalia')
lookup = data.frame(from_c, to_c)
happy$Country.name <- as.character(happy$Country.name)
index_to_replace = which(!is.na(match(happy$Country.name, lookup$from_c)))

# replacing country name
happy$Country.name[index_to_replace] = as.character(lookup$to_c[match(happy$Country.name,
                                                                     lookup$from_c)][index_to_replace])
```

#Rejoin table
```{r}
world$Country.Name = as.character(world$Country.Name)
new_table = left_join(x=happy, y=world, by=c("Country.name" = "Country.Name") )
unique(new_table[is.na(new_table['Land area (sq. km)']),][['Country.name']])
```


```{r}
str(countries)
```

# Just to retrieve the region information
```{r}
countries <- countries[c('Country', 'Region')]
```


# Merging happiness and countries dataset
# first to remove the stupid trailing whitespace in the country name
```{r}
levels(countries$Country) = sapply(levels(countries$Country), trimws, which="right")
new_table2 = left_join(x=new_table, y=countries, by=c("Country.name" = "Country") )
head(new_table2)
```

# Examine the new table - countries that are not matched
```{r}
unique(new_table2[is.na(new_table2$Region),][['Country.name']])
```

# Need to change the names manually
Bosnia and Herzegovina -> Bosnia & Herzegovina
Central Africa Republic -> Central African Rep.
Congo, Rep. -> Congo, Repub. of the 
Egypt, Arab Rep. -> Egypt 
Hong Kong SAR, China -> Hong Kong
Iran, Islamic Rep. -> Iran
Kosovo -> ****NO DATA****
Kyrgyz Republic -> Kyrgyzstan 
Lao PDR -> Laos
North Macedonia -> Macedonia
Montenegro -> ****NO DATA****
Myanmar -> Burma
West Bank and Gaza -> West Bank
Russian Federation -> Russia
Slovak Republic -> Slovakia
Korea, Rep. -> Korea, South
South Sudan -> Sudan
Eswatini -> Swaziland
Syrian Arab Republic -> Syria
Taiwan Province of China -> Taiwan
Trinidad and Tobago -> Trinidad & Tobago
Venezuela, RB -> Venezuela 
Yemen, Rep. -> Yemen


```{r}
## in happy dataset
from_c = c('Bosnia and Herzegovina', 'Central African Republic', 'Congo, Rep.' ,
           'Egypt, Arab Rep.', 'Hong Kong SAR, China', 'Iran, Islamic Rep.', 'Kyrgyz Republic',
           'Lao PDR', 'North Macedonia', 'Myanmar', 'West Bank and Gaza', 'Russian Federation',
           'Slovak Republic', 'Korea, Rep.', 'South Sudan', 'Eswatini', 'Syrian Arab Republic',
           'Taiwan Province of China', 'Trinidad and Tobago', 'Venezuela, RB', 'Yemen, Rep.')
## in countries dataset
to_c = c('Bosnia & Herzegovina', 'Central African Rep.', 'Congo, Repub. of the', 
         'Egypt', 'Hong Kong', 'Iran', 'Kyrgyzstan',
         'Laos', 'Macedonia', 'Burma', 'West Bank', 'Russia',
         'Slovakia', 'Korea, South', 'Sudan', 'Swaziland', 'Syria',
         'Taiwan', 'Trinidad & Tobago', 'Venezuela', 'Yemen')
lookup = data.frame(from_c, to_c)
new_table$Country.name <- as.character(new_table$Country.name)
index_to_replace = which(!is.na(match(new_table$Country.name, lookup$from_c)))

# replacing country names from new_table
new_table$Country.name[index_to_replace] = as.character(lookup$to_c[match(new_table$Country.name,
                                                                     lookup$from_c)][index_to_replace])
```


```{r}
unique(new_table$Country.name)
```


#Rejoin table
```{r}
countries$Country = as.character(countries$Country)
new_table2 = left_join(x=new_table, y=countries, by=c("Country.name" = "Country") )
unique(new_table2[is.na(new_table2$Region),][['Country.name']])
```


# Two countries without region = "Kosovo and Montenegro"
```{r}
str(new_table2)
```

```{r}
write_csv(new_table2, 'merged_dataset.csv')
```

